---
title: "feb18"
output: html_document
---
```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(blscrapeR)
```

1. Modify code to do 10+ states and as many years as possible
```{r}
for (state in c("AZ", "CA",  "CO", "DE", "MD", "FL", "DC", "UT", "NC", "WI")) {
   for (year in c(12:18)){
      bridges_URL <- paste("https://www.fhwa.dot.gov/bridge/nbi/20",
      year,
      "/delimited/",
      state,
      year,
      ".txt", sep = "")
   data <- read_delim(bridges_URL, ",", col_types = cols_only(
                                                      'STATE_CODE_001'= col_integer(),
                                                      'COUNTY_CODE_003' = col_integer(),
                                                      'STRUCTURE_NUMBER_008' = 'c',
                                                      'SERVICE_LEVEL_005C' = 'c',
                                                      'HIGHWAY_DISTRICT_002' = 'c', 
                                                      'COUNTY_CODE_003' = col_integer(), 
                                                      'MAINTENANCE_021' = col_integer(),
                                                      'YEAR_BUILT_027' = col_integer(),
                                                      'TRAFFIC_LANES_ON_028A' = col_integer(),
                                                      'DECK_COND_058' = 'c',
                                                      'SUPERSTRUCTURE_COND_059' = 'c',
                                                      'ADT_029' = col_integer()))
   data$DATA_YEAR <- paste("20", year, sep = "")
   nam <- paste("Bridges", state, year, sep = "_")
   assign(nam, data)
   }
}
```

```{r}
Bridges_AZ <- rbind.data.frame(Bridges_AZ_12, Bridges_AZ_13, Bridges_AZ_14, Bridges_AZ_15, Bridges_AZ_16, Bridges_AZ_17, Bridges_AZ_18)
Bridges_CA <- rbind.data.frame(Bridges_CA_12, Bridges_CA_13, Bridges_CA_14, Bridges_CA_15, Bridges_CA_16, Bridges_CA_17, Bridges_CA_18)
Bridges_CO <- rbind.data.frame(Bridges_CO_12, Bridges_CO_13, Bridges_CO_14, Bridges_CO_15, Bridges_CO_16, Bridges_CO_17, Bridges_CO_18)
Bridges_DE <- rbind.data.frame(Bridges_DE_12, Bridges_DE_13, Bridges_DE_14, Bridges_DE_15, Bridges_DE_16, Bridges_DE_17, Bridges_DE_18)
Bridges_MD <- rbind.data.frame(Bridges_MD_12, Bridges_MD_13, Bridges_MD_14, Bridges_MD_15, Bridges_MD_16, Bridges_MD_17, Bridges_MD_18)
Bridges_FL <- rbind.data.frame(Bridges_FL_12, Bridges_FL_13, Bridges_FL_14, Bridges_FL_15, Bridges_FL_16, Bridges_FL_17, Bridges_FL_18)
Bridges_DC <- rbind.data.frame(Bridges_DC_12, Bridges_DC_13, Bridges_DC_14, Bridges_DC_15, Bridges_DC_16, Bridges_DC_17, Bridges_DC_18)
Bridges_UT <- rbind.data.frame(Bridges_UT_12, Bridges_UT_13, Bridges_UT_14, Bridges_UT_15, Bridges_UT_16, Bridges_UT_17, Bridges_UT_18)
Bridges_NC <- rbind.data.frame(Bridges_NC_12, Bridges_NC_13, Bridges_NC_14, Bridges_NC_15, Bridges_NC_16, Bridges_NC_17, Bridges_NC_18)
Bridges_WI <- rbind.data.frame(Bridges_WI_12, Bridges_WI_13, Bridges_WI_14, Bridges_WI_15, Bridges_WI_16, Bridges_WI_17, Bridges_WI_18)

Bridges_Full <- rbind.data.frame(Bridges_AZ, Bridges_CA, Bridges_CO, Bridges_DE, Bridges_MD, Bridges_FL,
                                 Bridges_DC, Bridges_UT, Bridges_NC, Bridges_WI)
Bridges_Full$SUPERSTRUCTURE_COND_059 <- ifelse(Bridges_Full$SUPERSTRUCTURE_COND_059 == "N", NA, Bridges_Full$SUPERSTRUCTURE_COND_059)
Bridges_Full$STATE_CODE_001 <- as.factor(as.character(Bridges_Full$STATE_CODE_001))
Bridges_Full$SUPERSTRUCTURE_COND_059 <- as.numeric(as.character(Bridges_Full$SUPERSTRUCTURE_COND_059))
```

2. Explore more with ggplot
```{r}
#Make facets over states.
ggplot(data = subset(Bridges_Full, !is.na(SUPERSTRUCTURE_COND_059)), aes(x=YEAR_BUILT_027, y=SUPERSTRUCTURE_COND_059)) + geom_point() + geom_jitter() + facet_wrap(~STATE_CODE_001) + ggtitle("Superstructure Condition vs. Year Built")
#Fit linear model with bridges and unemployment data
#Need to import November/December data to predict future periods
```

```{r}
#Make one plot with x-axis as time and y-axis as something about bridges. make one line for each state.
#Going to make summary statistics to track change over time for the data
Bridges_Summarized <- Bridges_Full %>% 
   group_by(STATE_CODE_001, DATA_YEAR) %>% 
   summarize(Mean_Superstructure_Cond = mean(SUPERSTRUCTURE_COND_059, na.rm = T))
   
ggplot(Bridges_Summarized, aes(x=DATA_YEAR, y=Mean_Superstructure_Cond, group = STATE_CODE_001, color = STATE_CODE_001)) + geom_line() + ggtitle("Mean Superstructure Condition Per State")

#This graph shows Florida consistently has the highest mean superstructure condition, while Maryland has one of the lowest
```

Use the bridges data to find features that you think would be predictive of unemployment. Fit a linear model to predict the number of unemployed. Fit another model to predict the unemployment rate. Then, use the unemployed number and rate from the previous month as additional predictors. How do things change? Make a nice github repo with your work!
```{r}
# Grab BLS Unemployment Data To Start
BLS_data <- get_bls_county(c("September 2019", "December 2019"),
                           stateName = c("Arizona", "California", "Colorado", "Delaware", "Maryland", 
                                         "Florida", "District of Columbia", "Utah", "North Carolina",  "Wisconsin"))

```

```{r}
#Prepare Bridges data to be merged
Bridges_Full$STATE_CODE_001 <- as.numeric(as.character(Bridges_Full$STATE_CODE_001))
Bridges_Full$fips <- paste(Bridges_Full$STATE_CODE_001, Bridges_Full$COUNTY_CODE_003, sep ="")
Bridges_Full$fips <- as.numeric(as.character(Bridges_Full$fips))
BLS_data$fips <- as.numeric(as.character(BLS_data$fips))
Bridges_Full$DECK_COND_058 <- as.numeric(as.character(Bridges_Full$DECK_COND_058))

Bridges_of_interest <- Bridges_Full %>% 
   filter(DATA_YEAR == "2018") %>% 
   group_by(fips) %>% 
   summarize(mean_traffic_lanes = mean(TRAFFIC_LANES_ON_028A),
             mean_traffic = mean(ADT_029), 
             mean_deck_cond = mean(DECK_COND_058, na.rm = T),
             mean_superstructure = mean(SUPERSTRUCTURE_COND_059, na.rm = T))
```

```{r}
#Join Data
Bridges_and_BLS <- BLS_data %>% 
   inner_join(Bridges_of_interest, by = "fips")
```

```{r}
lm_unemployed <- lm(unemployed ~ mean_traffic + mean_deck_cond + mean_superstructure, data = Bridges_and_BLS)

summary(lm_unemployed)
#We see superstructure condition is not statistically significant
#Adj R-Sq of 0.6429
```
```{r}
lm_unemployment_rate <- lm(unemployed_rate ~ mean_traffic + mean_deck_cond + mean_superstructure, data = Bridges_and_BLS)

summary(lm_unemployment_rate)
#none of the predictors are statistically significant
#Adj R-Sq of 0.003431
```

```{r}
#Create lagged variable column to capture previous months BLS data
BLS_data_September2019 <- get_bls_county("September 2019",
                           stateName = c("Arizona", "California", "Colorado", "Delaware", "Maryland", 
                                         "Florida", "District of Columbia", "Utah", "North Carolina",  "Wisconsin"))
                           
BLS_data_December2019 <- get_bls_county("December 2019",
                           stateName = c("Arizona", "California", "Colorado", "Delaware", "Maryland", 
                                         "Florida", "District of Columbia", "Utah", "North Carolina",  "Wisconsin"))
BLS_data_September2019 <- BLS_data_September2019[, c(7,8), ]
colnames(BLS_data_September2019)[1] <- "unemployed_lag"
colnames(BLS_data_September2019)[2] <- "unemployment_rate_lag"

BLS_data <- cbind.data.frame(BLS_data_December2019, BLS_data_September2019)

BLS_data$fips <- as.numeric(as.character(BLS_data$fips))
#Join Data
Bridges_and_BLS <- BLS_data %>% 
   inner_join(Bridges_of_interest, by = "fips")
```

```{r}
#lagged regressor model
lm_unemployed <- lm(unemployed ~ mean_traffic + mean_deck_cond + mean_superstructure + unemployed_lag, data = Bridges_and_BLS)

summary(lm_unemployed)
#Unemployed lag is highliy significant
#Adj R-Sq of 0.8551, which is significantly higher than without lagged regressor from previous period
```

```{r}
#lagged regressor model
lm_unemployment_rate <- lm(unemployed_rate ~ mean_traffic + mean_deck_cond + mean_superstructure + unemployment_rate_lag, data = Bridges_and_BLS)

summary(lm_unemployment_rate)
#Unemployment_rate lag is highliy significant
#Adj R-Sq of 0.005158, which is higher, but still not good
```

Based on these models, it's clear that it's much easier to predict the number of unemployed persons in each fips county than the unemployment rate. The R-squared value for all of the unemployed models was drastically higher than the models that used the rate instead. Additionally, the lagged variables were highly significant, which shows us that the previous periods unemployed and unemployed_rate are both good predictors for the next period. Overall, it seems like bridges data gives us some insight into the number of unemployed persons in a fip code, but it is not perfect and could be improved. 